\section{Overview of Modeling Components}
\label{sec:ModelingComponents}

We consider a capital budgeting problem for a nuclear generation station, with possible
extension to a larger fleet of plants. Due to limited resources, we can only select a
subset from a number of candidate investment projects. Our goal is to maximize overall net
present value (NPV), or a variant of this objective, when we incorporate uncertainty into
project cost and revenue streams. In doing so, we must respect resource limits
and capture key structural and stochastic dependencies of the system. Example projects
include upgrading a steam turbine, refurbishing or replacing a set of reactor coolant pumps,
and replacing a set of feed-water heaters. Selecting an individual project has multiple
facets and implications.

\begin{itemize}
  \item \textbf{Rewards or Net Present Values}: Selecting a project can improve revenue (e.g.,
  upgrading a steam turbine may lead to an uprate in plant capacity resulting in larger
  revenue from selling power.) Replacing a key system component can improve reliability,
  increasing revenue due to a reduction in forced outages as well as operations and
  maintenance (O\&M) costs. Choosing to perform minimum maintenance versus refurbishing
  a component or replacing and improving a system can produce reward streams that
  can be negative or positive depending on the selection. Parameter
  \xmlNode{net\_present\_values} is used to specify the rewards (see Section~\ref{subsec:Parameters}).

  \item \textbf{Resources and Liabilities}: Critical resources, including (i) capital costs,
  (ii) O\&M costs, (iii) time and labor-hours during a planned outage, and (iv) personnel,
  installation and maintenance of equipment, workspaces, etc.. Within these categories, resources
  can further subcategorized, (witch each subcategory having its own budget), according to the plant’s organizational
  structure to provide multiple “colors” of money within capital costs, O\&M costs,
  personnel availability, etc.. The set \xmlNode{resources} and parameter
  \xmlNode{available\_capitals} are used to specify the resources and
  liabilities (see Section~\ref{subsec:Sets} and Section~\ref{subsec:Parameters}).

  \item \textbf{Costs}: Selecting a project in year $t$ induces multiple
  cost streams in year $t$ as well as in subsequent years; we interpret “cost” broadly to
  include commitment of critical resources. The parameter \xmlNode{costs} is used to specify
  the costs (see Section~\ref{subsec:Parameters}).

  \item \textbf{Time Periods}: Multiple capital projects can compete for same time period,
  limiting project selection. The set \xmlNode{time\_periods} is used to provide indices for
  \textbf{costs} and \textbf{available capitals} (see Section~\ref{subsec:Sets}).

  \item \textbf{Options}: The goal of selecting a project is typically to improve or maintain
  a particular function the plant performs, and there may be multiple ways to carry out
  the task. A project may be performed over a three-year period--say, years `$t$, $t+1$, $t+2$'--or the
  start of the project could instead be two years hence, changing the equation to
  `$t+2$, $t+3$, $t+4$'. Alternatively, at increased cost and benefit, it may be
  possible to complete the project in two years: `$t$, $t+1$' or `$t+2$, $t+3$'. When selecting a project
  to uprate plant capacity, we may have the options of increasing it by 3\% or 6\%.
  In all these cases, we can perform the project in, at most, one particular way, out of a collection of
  options. We represent this by cloning a “project” into multiple project-option pairs,
  and adding a constraint saying that we can select, at most, one from this set of options.
  The set \xmlNode{options} is used to provided indices for these multiple project-option pairs
  (see Section~\ref{subsec:Sets}).

  \item \textbf{Capitals}: If we consider maintenance for multiple units in an NPP in parallel,
  it has to be decided whether to accept a particular replacement and, in the positive
  case in which unit to conduct the corresponding replacement. In this case, the set \xmlNode{capitals}
  is used to provided indices for these units (see Section~\ref{subsec:Sets}).

  \item \textbf{Available Capitals}: This is available budgets for resources/units. The parameter
  \xmlNode{available\_capitals} is used to specify the available capitals for different
  resources/units at different $t$ (see Section~\ref{subsec:Parameters}).

  \item \textbf{Non-Selection}: Not selecting a project also has implications, inducing a growth
  in O\&M costs in future years, a decrease in plant production, an increase in forced outages,
  and even risking a premature end to plant life. Thus, not selecting a project can be seen as
  one more “option” for how a larger project is executed, expanding the list discussed earlier.
  Selection is of the “do nothing” option is reflected in both liability streams and reward
  streams. This can be activated by setting \xmlNode{nonSelection} to \xmlString{True}
  (see Section~\ref{subsec:Settings}).

  \item \textbf{Uncertainty}: One limitation of traditional optimization models for capital
  budgeting is that they do not account for uncertainty in reward and cost streams associated
  with individual projects, nor do they account for uncertainty in resource availability in
  future years. Projects can incur cost over-runs, especially when projects are large, performed
  infrequently, or when there is uncertainty regarding technical viability, external contractors,
  and/or suppliers of requisite parts and materials. Occasionally, projects are performed ahead
  of schedule and with savings in cost. Planned budgets for capital improvements can be cut, and key
  personnel may be lost. Or, there may be surprise budgetary windfalls for maintenance activities
  due to decreased costs for “unplanned” maintenance. The XML node \xmlNode{Uncertainties} is used
  to specify such uncertainties (see Section~\ref{subsec:Uncertainties}).

  %\item \textbf{Synergies}: Selecting a project may require replacing a structure, system, or
  %component (SSC) during a planned outage of the plant. Depending on the physical location of
  %an SSC in the plant and its relationship to other components, selecting one project may
  %reduce the cost of selecting another project (e.g., time or know-how required to implement
  %the project) if they are selected at the same time or close in proximity. For example, if
  %a plant has two units, selecting a project for one unit in a spring outage (e.g., replacement
  %of a condensate cooler and a set of feed-water heaters) may be followed by the same activity
  %in the fall outage in the second unit, at reduced cost.

  %\item \textbf{Planned Outage}: Nuclear power plants have planned outages at regular
  %intervals (e.g., every 18 months) often in the fall and spring to be well-prepared for winter
  %and summer peaks in load. While refueling only takes a fraction of a two-month (say) period
  %without power production, maintenance projects may be deferred until an outage. Moreover,
  %an outage can provide the only possible time period in which to carry out certain types of
  %projects. Because of lost revenue, an operator seeks to limit downtime. As a result, this
  %provides a special type of resource constraint limiting project selection due to multiple
  %projects competing for time, space, personnel, and equipment during an outage.

\end{itemize}

LOGOS consists of a collection of modeling entities/components that define different
aspects of the model, including \xmlNode{Sets}, \xmlNode{Parameters},
\xmlNode{Uncertainties}, and \xmlNode{ExternalConstraints}. In addition, the \xmlNode{Setting}
block specifies how the overall computation should run.

%
\subsection{Sets}
\label{subsec:Sets}

This subsection contains information regarding the XML nodes used to define the
\xmlNode{Sets} of the optimization model being performed through LOGOS.
\xmlNode{Sets} specifies a collection of data, possibly including
numeric data (e.g. real or integer values) as well as symbolic data (e.g. strings)
typically used to specify the valid indices for indexed components.
\nb Numeric data provided in \xmlNode{Sets} would be treated as strings.
\xmlNode{Sets} accepts the following additional sub-nodes:
\begin{itemize}
  \item \xmlNode{investments}, \xmlDesc{comma/space-separated string, required}, specifies
  the valid indices for investment projects.
  \item \xmlNode{capitals}, \xmlDesc{comma/space-separated string, optional},
  specifies the indices for NPP units.
  \item \xmlNode{time\_periods}, \xmlDesc{comma/space-separated string, optional},
  specifies the indices for time.
  \item \xmlNode{resources}, \xmlDesc{comma/space-separated string, optional},
  specifies indices for the resources and liabilities.
  \item \xmlNode{options}, \xmlDesc{semi-colon separated list of strings, optional},
  specifies the indices for multiple project-option pairs.
  This sub-node accepts the following attribute:
  \begin{itemize}
    \item \xmlAttr{index}, \xmlDesc{string, required}, specifies the index dependence.
    Valid index is \xmlString{investments}.
  \end{itemize}
\end{itemize}

Example XML:
\begin{lstlisting}[style=XML]
<Sets>
  <investments>
      HPFeedwaterHeaterUpgrade
      PresurizerReplacement
      ...
      ReplaceMoistureSeparatorReheater
  </investments>
  <time_periods>year1 year2 year3 year4 year5</time_periods>
  <resources>CapitalFunds OandMFunds</resources>
  <options index='investments'>
    PlanA PlanB DoNothing;
    PlanA PlanB PlanC;
    ...
    PlanA PlanB PlanC DoNothing
  </options>
</Sets>
\end{lstlisting}


%
\subsection{Parameters}
\label{subsec:Parameters}
This subsection contains information regarding the XML nodes used to define the
\xmlNode{Parameters} of the optimization model being performed through LOGOS:
\begin{itemize}
  \item \xmlNode{net\_present\_values}, \xmlDesc{comma/space-separated string, required},
  specifies the NPVs for capital projects or project-option pairs. This node accepts the
  following optional attribute:
  \begin{itemize}
    \item \xmlAttr{index}, \xmlDesc{comma-separated string, optional},
    specifies the indices of this parameter; keywords should be predefined in \xmlNode{Sets}.
    Valid keywords are \xmlString{investments} and \xmlString{options}.
    \default{investments}
  \end{itemize}
  \item \xmlNode{costs}, \xmlDesc{comma/space-separated string, required},
  specifies the costs for capital projects or project-option pairs. This node accepts the
  following optional attribute:
  \begin{itemize}
    \item \xmlAttr{index}, \xmlDesc{comma-separated string, optional},
    specifies the indices of this parameter; keywords should be predefined in \xmlNode{Sets}.
    Valid keywords are \xmlString{investments}, \xmlString{investments, time\_periods},
    \xmlString{options}, \xmlString{options, resources}, \xmlString{options, time\_periods},
    and \xmlString{options, resources, time\_periods}.
    \default{`investments'}
  \end{itemize}
  \item \xmlNode{available\_capitals}, \xmlDesc{comma/space-separated string, required},
  specifies the available capitals for capital projects or project-option pairs.
  This node accepts the following optional attribute:
  \begin{itemize}
    \item \xmlAttr{index}, \xmlDesc{comma-separated string, optional},
    specifies the indices of this parameter; keywords should be predefined in \xmlNode{Sets}.
    Valid keywords are \xmlString{resources}, \xmlString{time\_periods}, \xmlString{capitals},
    \xmlString{resources, time\_periods}, and \xmlString{capitals, time\_periods}
    \default{None}
  \end{itemize}
\end{itemize}

Example XML:
\begin{lstlisting}[style=XML]
<Parameters>
  <net_present_values index='options'>
    27.98 27.17 0.
    -10.07 -9.78 -9.22
    ...
    8.26 7.56 7.34 0.
  </net_present_values>
  <costs index='options,resources,time_periods'>
    12.99 1.3 0 0 0
    ...
    0.01 0 0 0 0
  </costs>
  <available_capitals index="resources,time_periods">
    22.6 36.7 20.6 23.6 22.7
    0.08 0.17 0.05 0.15 0.14
  </available_capitals>
</Parameters>
\end{lstlisting}


%
\subsection{Uncertainties}
\label{subsec:Uncertainties}
This subsection contains information regarding the XML nodes used to define the
\xmlNode{Uncertainties} of the optimization model being performed through LOGOS:
\begin{itemize}
  \item \xmlNode{available\_capitals}, \xmlDesc{optional}, specifies the scenarios
  associated with available capitals. This node accepts the attribute \xmlAttr{index}, which
  should be consistent with the \xmlNode{available\_capitals} defined in \xmlNode{Parameters}.
  This node accepts the following sub-nodes:
  \begin{itemize}
    \item \xmlNode{totalScenarios}, \xmlDesc{integer, required}, specifies the total
    number of scenarios for this parameter.
    \item \xmlNode{probabilities}, \xmlDesc{comma/space-separated float, required},
    specifies the probability for each scenario. The length should be equal to the total number of
    scenarios.
    \item \xmlNode{scenarios}, \xmlDesc{comma/space-separated float, required},
    specifies all scenarios for this parameter. The length should be equal to the total number
    of scenarios multiplied by the length of this parameter, as defined in \xmlNode{Parameters}.
  \end{itemize}

  \item \xmlNode{net\_present\_values}, \xmlDesc{optional}, specifies the scenarios
  associated with net\_present\_values. This node accepts the attribute \xmlAttr{index}, which
  should be consistent with the \xmlNode{net\_present\_values} defined in \xmlNode{Parameters}.
  \begin{itemize}
    \item \xmlNode{totalScenarios}, \xmlDesc{integer, required}, specifies the total
    number of scenarios for this parameter.
    \item \xmlNode{probabilities}, \xmlDesc{comma/space-separated float, required},
    specifies the probability for each scenario. The length should be equal to the total number of
    scenarios.
    \item \xmlNode{scenarios}, \xmlDesc{comma/space-separated float, required},
    specifies all scenarios for this parameter. The length should be equal to the total number
    of scenarios multiplied by the length of this parameter, as defined in \xmlNode{Parameters}.
  \end{itemize}
\end{itemize}

The overall number of scenarios is the total number of scenarios for \xmlNode{available\_capitals}
multiplied by the total number of scenarios for \xmlNode{net\_present\_values}.

Example XML:
\begin{lstlisting}[style=XML]
<Uncertainties>
  <available_capitals index="resources,time_periods">
    <totalScenarios>10</totalScenarios>
    <probabilities>
      0.5, 0.5
    </probabilities>
    <scenarios>
      20.0 34.0 17.0 20.0 18.0 0.08 0.17 0.05 0.15 0.14
      23.0 38.0 22.0 25.0 24.0 0.08 0.17 0.05 0.15 0.14
    </scenarios>
  </available_capitals>
  <net_present_values index='options'>
    <totalScenarios>9</totalScenarios>
    <probabilities>
      0.3 0.8
    </probabilities>
    <scenarios>
      13.3129 12.0228 0.0 -10.07
      ...
    </scenarios>
  </net_present_values>
</Uncertainties>
\end{lstlisting}


%
\subsection{External Constraints}
\label{subsec:ExternalConstraints}

This subsection contains information regarding the XML nodes used to define the
\xmlNode{ExternalConstraints} of the optimization model being performed through LOGOS.
This node accepts the following sub-node(s):
\begin{itemize}
  \item \xmlNode{constraint}, \xmlDesc{string, required}, specifies the external Python
  module file name along with its absolute or relative path. This external Python
  module contains the user-defined additional constraint.
  \nb If a relative path is specified, the code first checks relative to the working directory,
  then it checks with respect to the location of the input file. The working directory can be
  specified in \xmlNode{Settings} (see Section~\ref{subsec:Settings}). In addition, the extension
  `.py' is optional for the module file name that was inputted in this node.
  This sub-node also requires the following attribute:
  \begin{itemize}
    \item \xmlAttr{name}, \xmlDesc{string, required}, specifies the name of the constraint that will
    be added to the optimization problem.
  \end{itemize}
\end{itemize}

Example XML:
\begin{lstlisting}[style=XML]
<ExternalConstraints>
  <constraint name="con_I">externalConst</constraint>
  <constraint name="con_II">externalConstII.py</constraint>
</ExternalConstraints>
\end{lstlisting}

These constraints are Python modules, with a format automatically interpretable by
LOGOS. For example, users can define their own constraint, and the code will be embedded
and use the constraint as though it were an active part of the code itself.
The following provides an example of a user-defined external constraint:

Example Python Function:
\begin{lstlisting}[language=python]
# External constraint function
import numpy as np
import pyomo.environ as pyomo

def initialize():
  """
    Optional Method
    Optimization model parameters values can be updated/modified
    without directly accessing the optimization model.
    Value(s) will be updated in-place.
    @ In, None
    @ Out, updateDict, dict, {paramName:paramInfoDict},  where
      paramInfoDict contains {Indices:Values}
      Indices are parameter indices (either strings or tuples of
      strings, depending on whether there is one or
      more than one dimension). Values are the new values being
      assigned to the parameter at the given indices.
  """
  updateDict = {'available_capitals':{'None':16},
                'costs':{'1':1,'2':3,'3':7,'4':4,'5':8,
                         '6':9,'7':6,'8':10,'9':2,'10':5}
               }
  return updateDict

def constraint(var, sets, params):
  """
    Required Method
    External constraint provided by users that will be added to
    optimization problem
    @ In, sets, dict, all "Sets" provided in the Logos input
      file will be stored and available in this dictionary,
      i.e. {setName: setObject}
    @ In, params, dict, all "Parameters" provided in the
      Logos input file will be stored and
      available in this dictionary, i.e. {paramName: paramObject}
    @ In, var, object, the internally used decision variable,
      the dimensions/indices of this variable depend the type of
      optimization problems (i.e. "<problem_type>" from Logos
      input file). Currently, we will accept the following
      problem types:

      1. "singleknapsack": in this case, "var" will be var[:],
         where the index will be the element from
         xml node of "investment" in Logos input file.

      2. "multipleknapsack": in this case, "var" will be var[:,:],
         where the indices are the combinations element from set
         "investment" and element from set "capitals" in Logos
         input file

      3. "mckp": in this case, "var" will be var[:,:], where the
         indices are the combinations element from set
         "investment" and element from set "options" in Logos
         input file

      (Note that any element that is used as index will be
      converted to a string even if a number is provided in
      the Logos input file).

    @ Out, constraint, tuple, either (constraintRule,)
      or (constraintRule, indices)

    (Note that any modifications in provided sets and params
    will only have impact on this local module,
    i.e. the external constraint. In other words, the Sets
    and Params used in the internal constraints and
    objective will be kept unchanged!)
  """
  # All sets and parameters can be retrieved from dictionary
  # "sets" and "params" investments = sets['investments']

  def constraintRule(self, i):
    """
      Expression for user provided external constraint
      @ In, self, object, required to present, but not used
      @ In, i, str, element for the index set
      @ Out, constraintRule, function expression, expression
        to define user provided constraint

      Note that: Constraints can be indexed by lists or sets.
      When the return of function "constraint" contains
      lists or sets except the "constraintRule", the elements
      are iteratively passed to the rule function. If there
      is more than one, then the cross product is sent.
      For example, this constraint could be interpreted as
      placing limit on "ith" decision variable "var".
      A list of constraints for all "ith" decision variable
      "var" will be added to the optimization model
    """
    return var[i] <= 1

  # A tuple is required for the return, the first element
  # should be always the "constraintRule",
  # while the rest of elements are the lists or sets
  # if the user wants to construct the constraints
  # iteratively (See the docstring in "constraintRule"),
  # otherwise, keep it empty
  return (constraintRule, investments)
\end{lstlisting}

%
\subsection{Settings: Options for Optimization}
\label{subsec:Settings}

This subsection contains  information regarding the XML nodes used to define the
\xmlNode{Settings} of the optimization model being performed through LOGOS:
\begin{itemize}
  \item \xmlNode{problem\_type}, \xmlDesc{string, required parameter}, specifies the type of
  optimization problem. Available types include \xmlString{SingleKnapsack},
  \xmlString{MultipleKnapsack}, and \xmlString{MCKP} for risk-informed stochastic optimization.
  Available types include \xmlString{droskp}, \xmlString{dromkp}, and \xmlString{dromckp}
  for distributionally robust optimization.
  \item \xmlNode{solver}, \xmlDesc{string, optional parameter}, represents available solvers including
  \xmlNode{cbc} from \url{https://github.com/coin-or/Cbc.git} and \xmlNode{glpk} from
  \url{https://www.gnu.org/software/glpk/}.
  \item \xmlNode{sense}, \xmlDesc{string, optional parameter}, specifies \xmlString{minimize}
  or \xmlString{maximize} for minimization or maximization, respectively.
  \default{minimize}
  \item \xmlNode{mandatory}, \xmlDesc{comma/space-separated string, optional parameter},
  specifies regulatorily mandated or must-do projects.
  \item \xmlNode{nonSelection}, \xmlDesc{boolean, optional parameter}, indicates whether the
  investments options includes \textit{DoNothing} option.
  \default{False}
  \item \xmlNode{lowerBounds}, \xmlDesc{comma/space-separated integers, optional parameter}, specifies the lower bounds
  for decision variables.
  \item \xmlNode {upperBounds}, \xmlDesc{comma/space-separated integers, optional parameter}, specifies the upper bounds
  for decision variables.
  \item \xmlNode{consistentConstraintI}, \xmlDesc{string, optional parameter}, indicates whether
  this constraint is enabled.
  \default{True}
  \item \xmlNode{consistentConstraintII}, \xmlDesc{string, optional parameter}, indicates whether
  this constraint is enabled or not.
  \default{False}
  \item \xmlNode{solverOptions}, \xmlDesc{optional parameter}, will accept
  different options for the given solver provided in \xmlNode{solver}. A simple XML node only containing
  node tags and node texts can be used to provide the options for the solver. For example:
  \begin{lstlisting}[style=XML]
    <solverOptions>
      <threads>1</threads>
      <StochSolver>EF</StochSolver>
    </solverOptions>
  \end{lstlisting}
  In addition, if the problem type is distributionally robust optimization, additional option
  \xmlNode{radius\_ambiguity} can be used to control the Wasserstein distance. See Section~\ref{sec:DROCapitalBudgeting}.
\end{itemize}

Example XML:
\begin{lstlisting}[style=XML]
<Settings>
  <mandatory>
    PresurizerReplacement
    ...
    ReplaceInstrumentationAndControlCables
  </mandatory>
  <nonSelection>True</nonSelection>
  <consistentConstraintI>True</consistentConstraintI>
  <consistentConstraintII>True</consistentConstraintII>
  <solver>cbc</solver>
  <solverOptions>
    <threads>1</threads>
    <StochSolver>EF</StochSolver>
  </solverOptions>
  <sense>maximize</sense>
  <problem_type>mckp</problem_type>
</Settings>
\end{lstlisting}
